---
title: "Model_Selection_Of_Abalones"
author: "Brook Tarekegn Assefa"
date: "2022-11-14"
output:
  pdf_document: default
  html_document: default
---

## Model Selection Of Abalones


### Import the necessary packages 
```{r,echo=FALSE, include=FALSE}
library(leaps)
library(ggplot2)
library(tidyverse)
```

# Import the data and look at the first six rows

```{r}
Data <- read.csv("Data_Set--Abalone_Marine_Snails--With_Column_Names.csv")
head(Data)
```

# Let us first look at the Histogram of the rings Dataset, it seems to be slightly 
right skewed

```{r}
hist(Data$rings)
```
We might have to consider the Log transformation to flatten out the data and 
create a fairly Normal distribution ***Waiting for comment from other teammates***
```{r}
data<- Data %>% 
  select(-c(sex))
data<- log(data)
hist(data$rings)
```
### First, perform all possible regressions

The age and number of rings of a blacklip abalone has a moderate correlation 
with all variables other than sex and shucked weight of the abalone,
for which correlation is low.

***The following order was suggested in our proposal: shell weight, shucked***
***weight, diameter, whole weight, sex, viscera weight, and height. ***
As we can observer the if we consider one predictor model, 
what is the best one predictor?
This turns out to be the shell_weight as stated in our proposal.
If I consider two predictive model what turned out to be the best two predictive 
model are shell weight and shucked weight respectively; and the process goes on 
as per the suggested order.

```{r}
allregressions <- regsubsets(rings~., force.out=NULL, data=Data, nbest=1) 
summary(allregressions)
```
From this we would like to extract the best model based on the following criteria:-
**criteria 1**:- Having High/Maximum Adjusted R Squared value 
**criteria 2**:- Having Low Mallow's CP
**criteria 3**:- Having Low BIC 
The Mallow's Cp and BIC are usually similar. 
 
 ### These can now be extracted from the summary
```{r}
names(summary(allregressions))
```


### adjusted R 2 ?

```{r}
which.max(summary(allregressions)$adjr2)
```

### Mallow’s Cp ?

```{r}
which.min(summary(allregressions)$cp)
```

### BIC ?

```{r}
which.min(summary(allregressions)$bic)
```

Based on the above response of the *allregressions object* we can observe that 
Model 7 has the best adjusted R2 , Mallow's Cp and BIC. Therefore we can get
the corresponding coefficients and predictors of these models, and they all 
have shell_weight, sex, diameter, height, whole_weight, shucked_weight and 
viscera_weight.

Since Model 7 is the best model selected we can get the coefficient as follows:- 
```{r}
coef(allregressions, which.max(summary(allregressions)$adjr2))
```


We will then use the **Forward selection** , **Backward Selection** and 
the **Bidirectional selection** to find the best model according to AIC?


### Intercept only model
```{r}
regnull <- lm(rings~1, data=Data)
```


### Model with all predictors
```{r}
regfull <- lm(rings~., data=Data)
```

### Let us first carry out the  **Forward selection**.
```{r}
step(regnull, scope=list(lower=regnull, upper=regfull), direction="forward")
```
Therefore we can understand from the **Forward selection** all predictors are included 
except the **length predictor**. Indicating that the seven predictors listed below 
are significant predictors.

### Therefore the regression equation selected is as the following for y hat:- 

$$
\hat{y} = 3.87038 + 8.75078x_1 -19.80258x_2 + 10.56951 x_3 + 8.97751x_4 -0.76889x_5
-10.61279x_6 + 10.74911x_7
$$


### Let us now carry out the  **Backward selection**.
```{r}
step(regfull, scope=list(lower=regnull, upper=regfull), direction="backward")
```



Therefore we can understand from the **Backward selection** also includes all predictors 
except the **length predictor**. Indicating that the seven predictors listed below 
are significant predictors. T
### Therefore the regression equation selected is as the following for y hat:- 

$$
\hat{y} = 3.87038 + 8.75078x_1 -19.80258x_2 + 10.56951 x_3 + 8.97751x_4 -0.76889x_5
-10.61279x_6 + 10.74911x_7
$$

This doesn't help much as it is similar to the forward selection. 

Since in our proposal we have mentioned that the observed correlation among the 
rings has a moderate correlation with all variables other than
sex and shucked weight of the abalone. We would like to force out these values and perform the 
**Backward selection**.
```{r}
data_set_with_numeric_sex <-
    Data %>%
    mutate(sex = replace(sex, sex == 'M', 0)) %>%
    mutate(sex = replace(sex, sex == 'F', 1)) %>%
    mutate(sex = replace(sex, sex == 'I', 2))
data_set_with_numeric_sex$sex <- as.numeric(data_set_with_numeric_sex$sex)
correlation_matrix <- cor(data_set_with_numeric_sex)
correlation_matrix
```

### Model with all predictors except excluding the sex and shucked weight of the abalone.
```{r}
regfull_excluding_sex_shucked_weight <- lm(rings~ diameter+height+whole_weight+viscera_weight+shell_weight+length, data=Data)
```

### Perform the Backward elimination from the regfull

```{r}
step(regfull_excluding_sex_shucked_weight, scope=list(lower=regnull, upper=regfull_excluding_sex_shucked_weight), direction="backward")
```

After excluding the sex and shucked_weight we can observe that this change has rendered 
the **viscera weight** as insignificant predictor.
### Therefore our adjusted regression equation for y hat is :- 

$$
\hat{y} = 3.702 + 14.387x_1 + 13.085x_2 - 5.959x_3 + 26.202x_4 -5.308x_5
$$

 If We would like to force out the sex and shucked weight of the abalone 
 and perform the **Bidirectional selection**, we would have found the same result 
 as the one stated above.
 
 
But if we do perform the **Forward selection** we can see that the *viscera_weight*
weight is considered as a significant predictor , 
which was dropped by **Backward and Bidirectional selections.**
```{r}
step(regfull_excluding_sex_shucked_weight, scope=list(lower=regnull, upper=regfull_excluding_sex_shucked_weight), direction="forward")
```
 ### Therefore our Forward selection regression equation for y hat is :- 

$$
\hat{y} = 3.674 + 14.318x_1 + 13.176x_2 - 5.671x_3 -1.203x_4 + 26.017x_5 -5.183x_6  
$$
 
 
Therefore 
**Forward selection**:- 
$$
x_1 , x_2 , x_3 ,x_4 ,x_5 , x_6  
$$
**Backward elimination** and **Stepwise regression** 

$$
x_1 , x_2 , x_3 ,x_4 ,x_5
$$
"***Berk [1978]*** has noted that **forward selection**
tends to agree with all possible regressions for **small subset sizes** but not for large
ones, while b**ackward elimination** tends to agree with all possible regressions for
large subset sizes but not for small ones." 
Based on this our final regression equation and the model selection based on the AIC is 
excluding the  **viscera weight** (which is of a blacklip abalone is the gut weight after bleeding.)
When thinking logically about it , it seems like it does not have any relation
with the age of an abalone at all.
$$
\hat{y} = 3.702 + 14.387x_1 + 13.085x_2 - 5.959x_3 + 26.202x_4 -5.308x_5
$$
 
 